Article Title

### 2013-12-16 plos.tex
Rev23.  Incorporated some of Alex' comments.

\section*{Abstract}
% Please keep the abstract between 250 and 300 words
Ambitious projects currently under way aim to record the activity of ever larger and denser subsets of neurons \emph{in vivo}.  Correlations measured in such recordings are anticipated to reveal important aspects of the functional organization of neural circuits.  However, estimation and interpretation of large correlation matrices from finite recordings are challenging.  Estimation can be improved by regularization: biasing of the estimate toward a low-dimensional, less variable approximation.  The amount of improvement depends on how parsimoniously the reduced approximation captures the dominant dependencies in the data.  Therefore, the selection of the most efficient estimator is an empirical question that informs about the types of dominant dependencies present in the system.

In this study, we sought the most statistically efficient estimator of neural correlations in recordings from large, dense groups of cortical neurons.  Using fast 3D random-access laser scanning microscopy of calcium signals, we recorded the activity of nearly every neuron in volumes of about 100 $\mu$m in diameter (150--350 cells) in mouse visual cortex.  We hypothesized that, in these dense recordings, the correlation matrix would be most efficiently represented by a sparse network of linear interactions between pairs of observed neurons combined with common latent factors representing unobserved inputs and global network fluctuations.  Indeed, in cross-validation tests, the covariance matrix estimator that imposed this correlation structure outperformed estimators regularized toward other plausible low-dimensional approximations. Furthermore, we found consistent relationships between the organization of sparse interactions inferred by the estimator and the physical distances and preferred orientation differences of pairs of cells.  Consistent with previous synaptic connectivity studies, the density of positive interactions decreased rapidly with distance and with difference in preferred orientation whereas negative interactions were less selective.  To further corroborate physiological interpretation of the inferred functional structure, future experiments will augment this analysis with measured synaptic connectivity, cell types, and brain states.

\section*{Author Summary}
% Please keep the Author Summary between 150 and 200 words
% Use first person. PLoS ONE authors please skip this step. 
% Author Summary not valid for PLoS ONE submissions.   
Correlations of the activity of neurons have proven useful as descriptors of the functional organization of neural circuits with implications for stimulus coding and circuit architecture.  Estimation of correlation matrices can be improved by imposing some kind of structure on the estimate with greatest improvement attained when the imposed structure closely matches the real dependencies in the data. Using fast 3D two-photon imaging of calcium signals, we recorded the activity of large and dense groups of cells in mouse visual cortex and evaluated the cross-validated performance of correlation matrix estimators that imposed different kinds of structure. The correlation structure of the estimator that proved most efficient comprised a sparse network of partial correlations between pairs of neurons combined with several common latent factors, with both components required for efficient estimation. Because of its superior benefit for estimation, we proposed that this inferred structure could prove more relevant for the description of functional connectivity than the usual correlation matrix in densely sampled neural recordings. As a first application of this approach, we analyzed how the inferred connectivity related to distances between cells and differences in their preferred orientations and found basic agreement with previous studies of synaptic connectivity.

\section*{Introduction}
\subsection*{Neural correlations}
Pearson correlations between the spiking activity of pairs of neurons, or simply \emph{neural correlations}, are the most familiar descriptive statistics of neural population activity \cite{Averbeck:2006,Zohary:1994,Kohn:2005,Bair:2001,Ecker:2010}.  For example, \emph{noise correlations}, \emph{i.e.}~the correlations of stimulus response variability between pairs of neurons, have been shown to have profound theoretical implications for stimulus coding \cite{Zohary:1994,Abbott:1999,Averbeck:2006,Berens:2011,Ecker:2011}. In addition, neural correlations are hypothesized to reflect key features of functional connectivity in neural circuits.  Such interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008,Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus tuning similarity \cite{Kohn:2005,Ko:2011}, cortical layer specificity \cite{Hansen:2012,Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009}, changes due to sensory stimulation and global brain states \cite{Goard:2009,Kohn:2009,Ecker:2010,Renart:2010}, and others.

@@

For example, one form of low-dimensional representation of the correlation matrix is its low-rank approximation. Low-rank approximations are particularly suitable for capturing shared fluctuations across the recorded population. Extracted through principal component or factor analysis, low-rank components of neural correlations have been analyzed in a number of studies with particular attention paid to their temporal dynamics \cite{Yu:2006,Yu:2009}.

Alternatively, the correlation matrix can be formulated through the partial correlations between pairs of neurons. The partial correlation between a pair of neurons is the linear correlation remaining after accounting for correlations to all the other neurons. Partial correlations carry particular significance when conditional dependencies between all the variables are thought to be dominated by linear effects. In such systems, zero  partial correlation between two variables implies their conditional independence.  Conditional independence between a pair of variables suggests a lack of direct interaction between the underlying processes. Therefore, estimation of networks of partial correlations (sometimes called \emph{association networks}) have been used to infer interactions between components in complex systems such as gene interaction networks \cite{Schafer:2005,Peng:2009} and functional connectivity between brain regions in fMRI signals \cite{Varoquaux:2012,Ryali:2012}. In \emph{elliptical distributions} (those of the form $\frac 1 Z g\left((x-\mu)^\T\Sigma^{-1}(x-\mu)\right)$, where $g$ is an arbitrary function, $Z$ is a normalizer, $\Sigma$ is the covariance matrix) and the multivariate normal distribution in particular, interactions are completely defined by linear effects and partial interactions express conditional dependencies. A multivariate normal distribution specified through its graph of non-zero partial correlations is known as a Gaussian Graphical Model or Gauss-Markov Random Field \cite{Koller:2009}. With increasing departure from linearity, the correspondence between conditional dependence and partial correlations diminishes or completely breaks down \cite{Loh:2012}. For example, the coupling terms in pairwise Ising models have no direct relationship to the partial correlations.

@@

Given a covariance matrix estimate $C$, the correlation matrix $R$ and partial correlation matrix $P$ are computed as
\begin{equation}
    R = \left(I\circ C\right)^{-\frac 1 2} C \left(I\circ C\right)^{-\frac 1 2}
\end{equation}
\begin{equation}
    P = -\left(I\circ C^{-1}\right)^{-\frac 1 2} C^{-1} \left(I\circ C^{-1}\right)^{-\frac 1 2}
\end{equation}
Here $\circ$ denotes entrywise matrix product (Hadamard product) and $I$ is the $p\times p$ identity matrix. $R$ contains 1s on its diagonal whereas $P$ contains $-1$s on the diagonal. The off-diagonal elements of $P$ contain the partial correlations between corresponding pairs of neurons.

@@

Estimation can be improved through \emph{regularization}: the deliberate biasing of the estimate toward a low-dimensional approximation \cite{Schafer:2005,Bickel:2006}.  The usual estimate $C_{\sf 0}$ has the advantage of being unbiased ($\E{C_0}=\Sigma$) but, on average, it falls far from $\Sigma$ due to its sensitivity to sampling noise in the data.  Low-dimensional estimates of various forms are typically less susceptible to sampling noise but are also liable to introduce their respective biases away from truth $\Sigma$.  Regularization works by striking a favorable balance between bias and variability. Regularization can produce \emph{some} improvement even with an arbitrary target estimate that has no relation to the true dependencies in the data.  Yet, when the target estimate is well suited for capturing the important features of the true covariance matrix with few terms, it will introduce minimal bias and outperform other estimators.

Figure \ref{fig:01} illustrates the challenges of estimating the correlation matrix from a finite sample and one example of regularization.  Panel A depicts the correlation matrix computed from the sample covariance matrix $C_{\sf 0}$ for a period of somatic calcium signals of a group of 298 neurons in mouse visual cortex. To show detail, panel B magnifies the fragment outlined in panel A. Panel C shows the same fragment from the partial correlation matrix also computed from $C_{\sf 0}$.  Due to the low-pass filtration effect of calcium dye kinetics, correlations in unprocessed calcium signals are higher than typical firing rate correlations on shorter temporal scales.

Panels \ref{fig:01}\,D and E depict the correlation matrix and the partial correlation matrix computed from a different estimate, which we will refer to here as $C_{\sf sparse}$, which is computed through \emph{covariance selection} \cite{Dempster:1972}. In covariance selection, only a subset of partial correlations is fitted while the rest are fixed at zero. The resulting estimate closely reproduces the sample correlation matrix (Fig.~\ref{fig:01}\,D), but its partial correlation structure appears substantially different(Fig.~\ref{fig:01}\,E). In this particular example  31,501 of the 44,253 (71.2\%) total partial correlations are set to zero. This alternative estimate is attractive for a few reasons. First, it has fewer free parameters and is therefore less susceptible to sampling noise. Additionally, assuming that partial correlations do indeed reflect elements of the functional connectivity, the second estimate also appears more physiologically plausible: in neural circuits, each neuron only interacts with a fraction of the entire population.

\subsection*{Model selection}
Covariance selection is but one of numerous possible low-dimensional approximations of the covariance matrix. Indeed any probabilistic model of the joint population activity regularized toward a small number of parameters automatically defines a low-dimensional approximation of the covariance matrix. Models that accurately reflect the dominant functional dependencies in the circuit, its so-called ``functional connectivity'', will tend to produce estimates that are nearest to the true value of the covariance matrix and are said to be more \emph{efficient} or to \emph{dominate} other estimators. An estimator's superior efficiency can be ascertained by cross-validation.  Because different types of neural circuits are governed by different types of dependencies, different probabilistic models and covariance matrix estimators may dominate in each case. The estimator that is shown to dominate all others for a specific type of neural circuit may be proposed as the best descriptor of the low-dimensional representation of the correlation structure of the system of interest. Its structure may then be analyzed and related to the circuits anatomical organization.

\subsection*{Summary of findings}
In this study, we compared four regularized covariance matrix estimators biased toward different respective low-dimensional correlation structures: `shrinkage toward diagonal' or $C_{\sf diag}$, `shrinkage toward a multifactor model' or $C_{\sf factor}$, `sparse partial correlations' or $C_{\sf sparse}$, and 'sparse partial correlations with latent units' or $C_{\sf sparse+lowrank}$.  First, we demonstrated that, in simulations with known true low-dimensional correlation structures, regularized estimators with matching types of low-dimensional structure generally dominated over the other estimators. We then used cross-validation to determine the most efficient estimator for the population activity of dense groups of neurons in mouse primary visual cortex. We found that $C_{\sf sparse+lowrank}$ consistently dominated the other estimators.  Estimate produced by $C_{\sf sparse+lowrank}$ revealed networks of interactions that differed substantially from networks of strongest correlations and depended strongly on the physical distance separating pairs of cells and on the differences in their preferred orientations.

@@

Finally, we consider the fourth estimator $C_{\sf sparse+lowrank}$, which provides for both common latent factors interacting with all recorded neurons and sparse interactions between the recorded neurons (Fig.~\ref{fig:02}\,D). This estimator has the form \cite{Chandrasekaran:2010,Ma:2013}
\begin{equation}
C_{\sf sparse+latent} = (S - LL^\T)^{-1}
\end{equation}
where, as above, $S$ is a sparse matrix and $L$ is a $d\times p$ matrix of factor loadings. The estimator has two hyperparameters: the number of latent units $d$ and the sparsity of $S$.

@@

In simulation, CV-loss accurately reproduced the differences between the estimators' excess losses, although with greater variability (Figure \ref{fig:02}, Row 5). For each kind of ground truth (Row 2), relative CV-losses were computed with respect to the estimator whose regularization target matched the structure of the respective ground truth: $\ell_{C,C_{\sf diag}}$, $\ell_{C,C_{\sf factor}}$, $\ell_{C,C_{\sf sparse}}$, and $\ell_{C,C_{\sf sparse+lowrank}}$. Just as with excess loss in Row 4, the means and their standard errors were computed from 30 samples  taken for each ground truth and for each sample size.

@@

In these highly localized populations both direct interactions between cells and common diffuse inputs are likely to contribute to the overall population variability, leading us to hypothesize that regularized estimates capable of accommodating partial correlations between specific pairs of cells (\emph{e.g.}\;$C_{\sf sparse}$ and $C_{\sf sparse+lowrank}$) would be more efficient than others. At the same time, average correlations were relatively small (Fig.~\ref{fig:03}\,G), to the advantage of estimates biased toward independence (\emph{e.g.}\;$C_{\sf diag}$). Finally, there is the possibility that the correlations structure is best explained by common fluctuations across the entire population, to the advantage of estimates biased toward low-rank correlation structure (\emph{e.g.}\;$C_{\sf factor}$, $C_{\sf factor+lowrank}$).

We found that estimator $C_{\sf sparse+lowrank}$ was more efficient than the other estimators (Fig.~\ref{fig:04}). The median relative CV-validation loss (Eq.~\ref{eq:rel-cv-loss}) of each of the estimates $C_{\sf 0}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ with respect to $C_{\sf sparse+lowrank}$ was significantly above zero ($p<0.001$, Wilcoxon signed rank test, for each).

\subsection*{Relationship of correlation structure to circuit architecture}
Having demonstrated that $C_{\sf sparse+lowrank}$ dominated the other estimators, we examined the elements of the low-dimensional structure revealed by this estimator for individual imaged sites. By its design, the sparse+lowrank estimator finds the optimal balance between a sparse network of partial correlations and shared common latent units. If this estimate dominates all other estimators evaluated thus far, it seems reasonable to hypothesize that these interaction may reflect underlying physiological interactions more closely than the usual correlations. In particular, the sparse component of the partial correlation matrix suggests direct interactions between pairs of neurons, whereas the low-rank component suggest common fluctuations such as those caused by common inputs or other collective synchronized activity.

We examined the structure of the sparse+lowrank estimate in individual sites such as the example site first depicted in Fig.\;\ref{fig:03}. In this site, as in others, the regularized estimate of the correlation matrix appeared very similar to the unregularized sample correlation matrix (Fig.~\ref{fig:05}\,A and D). However, the corresponding partial correlation matrices differed dramatically (Fig.~\ref{fig:05}\,B and E). The partial correlation was decomposed into the sparse and low-rank components (Fig.\,\ref{fig:05}\,C). Although correlations were mostly positive, the sparse partial correlations (or `interactions'), although mostly positive, had a much larger proportion of negative values. The sparse component had 82.1\% sparsity (or 17.9\% connectivity), which corresponded to the average node degree (interactions per cell) of 52.5 (Fig.\;\ref{fig:05}\,G). The low-rank component was of rank 17.

In previous publications that looked at the structure of correlations, the sparse network was obtained by thresholding the correlations coefficients  above some level deemed significant and examining the network of correlations above that threshold \cite{Golshani:2009,Malmersjo:2013}. There was fairly little overlap between the network of interactions revealed by this thresholding method and those revealed by the sparse+lowrank estimator (Fig.\;\ref{fig:05}\,F). When thresholded to the same sparsity (82.1\%), only 42\% of cell pairs connected in one network were connected in the other while the average magnitude of such correlations was much lower in the case of the regularized estimator (Fig.\;\ref{fig:05}\,F, H, and I). In particular, many low sample correlations translated into negative interactions in the regularized estimate. Indeed, the absence of a correlation between pairs of cells that both correlate similarly to several of their neighbors should be considered as significant as a high correlation coefficient. Regularized partial correlations reveal such phenomena whereas regular sample correlations cannot.

@@

\begin{figure}[!ht]
    \begin{center}
        %\includegraphics[width=8.3cm]{Figure04.pdf}
    \end{center}
    \caption{{\bf The sparse+lowrank estimator $C_{\sf sparse+lowrank}$ outperforms the other estimators on neural data.}
    {\bf A--D.} Histograms of average cross-validation loss differences of the respective estimators $C_{\sf 0}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ from $C_{\sf sparse+lowrank}$. 
    The histograms are based on 31 imaged sites in 24 mice. 
    All medians (red dashed lines) were significantly greater than zero, indicating the dominance of $C_{\sf sparse+lowrank}$ over the other estimators. 
    The green arrows indicate the results for the site shown in Fig.~\ref{fig:03} and Fig.~\ref{fig:05}
    }
    \label{fig:04}
\end{figure}

>>>>

\section*{Abstract}
% Please keep the abstract between 250 and 300 words
Ambitious projects currently under way aim to record the activity of ever larger and denser subsets of neurons \emph{in vivo}.  Correlations measured in such recordings are anticipated to reveal important aspects of the functional organization of neural circuits.  However, estimation and interpretation of large correlation matrices from finite recordings are challenging.  Estimation can be improved by regularization: biasing of the estimate toward a low-dimensional, less variable approximation.  The amount of improvement depends on how parsimoniously the reduced approximation captures the dominant dependencies in the data.  Therefore, the selection of the most efficient estimator is an empirical question that informs about the types of dominant dependencies governing the system.

In this study, we sought the most statistically efficient estimator of neural correlations in recordings from large, dense groups of cortical neurons.  Using fast 3D random-access laser scanning microscopy of calcium signals, we recorded the activity of nearly every neuron in volumes of about 100 $\mu$m in diameter (150--350 cells) in mouse visual cortex.  We hypothesized that, in these dense recordings, the correlation matrix would be most efficiently represented by a sparse network of linear interactions between pairs of observed neurons combined with common latent factors representing unobserved inputs and global network fluctuations.  Indeed, in cross-validation tests, the covariance matrix estimator that imposed this correlation structure outperformed estimators regularized toward other plausible low-dimensional approximations. Furthermore, we found consistent relationships between the organization of sparse interactions inferred by this estimator and the physical distances and preferred orientation differences of pairs of cells.  Consistent with previous synaptic connectivity studies, the density of positive interactions decreased rapidly with distance and with difference in preferred orientation whereas negative interactions were less selective.  To further corroborate physiological interpretation of the inferred functional structure, future experiments will augment this analysis with measured synaptic connectivity, cell types, and brain states.

\section*{Author Summary}
% Please keep the Author Summary between 150 and 200 words
% Use first person. PLoS ONE authors please skip this step. 
% Author Summary not valid for PLoS ONE submissions.   
Correlations of the activity of neurons have proven useful as descriptors of the functional organization of neural circuits with implications for stimulus coding and circuit architecture.  Estimation of correlation matrices can be improved by imposing some kind of structure on the estimate with greatest improvement attained when the imposed structure efficiently captures real dependencies in the data. Using fast 3D two-photon imaging of calcium signals, we recorded the activity of large and dense groups of cells in mouse visual cortex and evaluated the cross-validated performance of correlation matrix estimators that imposed different kinds of structure. The correlation structure of the estimator that proved most efficient comprised a sparse network of partial correlations between pairs of neurons combined with several common latent factors, with both components required for efficient estimation. Because of its superior benefit for estimation, we proposed that this inferred structure could prove more relevant for the description of functional connectivity than the usual correlation matrix in densely sampled neural recordings. As a first application of this approach, we analyzed how the inferred connectivity related to distances between cells and differences in their preferred orientations and found basic agreement with previous studies of synaptic connectivity.

\section*{Introduction}
\subsection*{Neural correlations}
Pearson correlations between the spiking activity of pairs of neurons, or simply \emph{neural correlations}, are the most familiar descriptive statistics of neural population activity \cite{Averbeck:2006,Zohary:1994,Kohn:2005,Bair:2001,Ecker:2010}.  For example, \emph{noise correlations}, \emph{i.e.}~the correlations of stimulus response variability between pairs of neurons, have been shown to have profound theoretical implications for stimulus coding \cite{Zohary:1994,Abbott:1999,Sompolinsky:2001,Wilke:2002,Averbeck:2006,Josic:2009,Berens:2011,Ecker:2011}. In addition, neural correlations are hypothesized to reflect key features of functional connectivity in neural circuits.  Such interpretation is supported by a series of discoveries of nontrivial relationships between neural correlations and other aspects of circuit organization such as the physical distances between neurons \cite{Smith:2008,Denman:2013}, their synaptic connectivity \cite{Ko:2011},  stimulus response similarity \cite{Bair:2001,Kohn:2005,Cohen:2008,Cohen:2009,Ko:2011,Smith:2013b}, cortical layer specificity \cite{Hansen:2012,Smith:2013}, progressive changes in development and in learning \cite{Golshani:2009,Gu:2011}, changes due to sensory stimulation and global brain states \cite{Goard:2009,Kohn:2009,Ecker:2010,Renart:2010}, and others.

@@

For example, one form of low-dimensional representation of the correlation matrix is its low-rank approximation. Low-rank approximations are particularly suitable for capturing shared fluctuations across the recorded population. Extracted through principal component or factor analysis, low-rank components of neural correlations have been analyzed in a number of studies with particular attention paid to their temporal dynamics \cite{Yu:2009}.

Alternatively, the correlation matrix can be formulated through the partial correlations between pairs of neurons. The partial correlation between a pair of neurons is the linear correlation remaining after accounting for correlations to all the other neurons. Partial correlations carry particular significance when conditional dependencies between all the variables are thought to be dominated by linear effects. In such systems, zero  partial correlation between two variables implies their conditional independence.  Conditional independence between a pair of variables suggests a lack of direct interaction between the underlying processes. Therefore, estimation of networks of partial correlations (sometimes called \emph{association networks}) have been used to infer interactions between components in complex systems such as gene interaction networks \cite{Schafer:2005,Peng:2009} and functional connectivity between brain regions in fMRI signals \cite{Varoquaux:2012,Ryali:2012}. In \emph{elliptical distributions} (those of the form $x \sim \frac 1 Z g\left((x-\mu)^\T\Sigma^{-1}(x-\mu)\right)$, where $g$ is an arbitrary function, $Z$ is the scalar normalizer, $\Sigma$ is the covariance matrix) and the multivariate normal distribution in particular, interactions are completely defined by linear effects and partial interactions express conditional dependencies. A multivariate normal distribution specified through its graph of non-zero partial correlations is known as a Gaussian Graphical Model or Gauss-Markov Random Field \cite{Koller:2009}. With increasing departure from linearity, the correspondence between conditional dependence and partial correlations diminishes or completely breaks down \cite{Loh:2012}. For example, the coupling terms in pairwise Ising models have no direct relationship to the partial correlations \cite{Schneidman:2006,Tkacik:2006}.

@@

Given a covariance matrix estimate $C$ (not necessarily sample covariance $C_{\sf 0}$), the correlation matrix $R$ is calculated by normalizing $C$ by its diagonal (variance estimates):
\begin{equation}
    R = \left(I\circ C\right)^{-\frac 1 2} C \left(I\circ C\right)^{-\frac 1 2}
\end{equation}
Similarly, the matrix of partial correlations $P$ is computed as the negative normalized version of the \emph{precision matrix} (inverse covariance matrix) $C^{-1}$:
\begin{equation}
    P = -\left(I\circ C^{-1}\right)^{-\frac 1 2} C^{-1} \left(I\circ C^{-1}\right)^{-\frac 1 2}
\end{equation}
Here $\circ$ denotes entrywise matrix product (Hadamard product) and $I$ is the $p\times p$ identity matrix.

@@

Estimation can be improved through \emph{regularization}: the deliberate biasing of the estimate toward a low-dimensional approximation \cite{Schafer:2005,Bickel:2006}.  The usual estimate $C_{\sf 0}$ has the advantage of being unbiased ($\E{C_0}=\Sigma$) but, on average, it falls far from $\Sigma$ due to its sensitivity to sampling noise in training sample.  Low-dimensional estimates of various forms are typically less susceptible to sampling noise but are also liable to introduce their respective biases away from truth $\Sigma$.  Regularization works by striking a favorable balance between bias and variability. Regularization can produce \emph{some} improvement even with an arbitrary target estimate that has no relation to the true dependencies in the data.  Yet, when the target estimate is well suited for capturing the important features of the true covariance matrix with few terms, it will introduce minimal bias and outperform other estimators.

To illustrate challenges of estimating the correlation matrix from a finite sample, we briefly consider a regularization scheme based on \emph{covariance selection} (Figure \ref{fig:01}) \cite{Dempster:1972}. In covariance selection, the estimate is produced by fitting only an optimal subset of the coefficients of the precision matrix while setting the rest to zero. We estimated the covariance matrix of a period of a somatic calcium signals of a group of 298 neurons in mouse visual cortex using both the sample covariance matrix (Fig.\;\ref{fig:01}\,A, B) and covariance selection (Fig.\;\ref{fig:01}\,C). Due to the low-pass filtration effect of calcium dye kinetics, correlations in unprocessed calcium signals are higher than typical firing rate correlations on shorter temporal scales. The apparent similarity of the two estimates belies the utterly different partial correlation structure (Fig.\;\ref{fig:01}\,D, E): the regularized estimate is produced from the same data by zeroing the optimal set of 31,501 (71.2\%) of the off-diagonal coefficients of the precision matrix and fitting the remainder. Estimation by covariance selection is attractive for several reasons: First, it has fewer free parameters and is therefore less susceptible to sampling noise, yielding estimates that are, on average, closer to truth $\Sigma$. Additionally, if we assume that partial correlations do indeed reflect elements of the functional connectivity in the circuit, the second estimate also appears more plausible: in neural circuits, each neuron only interacts with a fraction of the entire population.

\subsection*{Model selection}
However, covariance selection is but one of numerous possible low-dimensional approximations of the covariance matrix. Indeed any probabilistic model of the joint population activity regularized toward a small number of parameters automatically defines a low-dimensional approximation of the covariance matrix. Models that accurately reflect the dominant functional dependencies in the circuit, its so-called `functional connectivity', will tend to produce estimates that are nearest to the true value of the covariance matrix and are said to be more \emph{efficient} than the other estimators or to \emph{dominate} the other estimators. An estimator's superior efficiency can be ascertained by cross-validation.  Because different types of neural circuits are governed by different types of dependencies, different probabilistic models and covariance matrix estimators may dominate in each case. The estimator that is shown to dominate all others for a specific type of neural circuit may be proposed as the best descriptor of the low-dimensional representation of the correlation structure of the system of interest. Its structure may then be analyzed and related to the circuits anatomical organization.

\subsection*{Summary of findings}
In this study, we compared four regularized covariance matrix estimators biased toward different respective low-dimensional correlation structures: `shrinkage toward diagonal' or $C_{\sf diag}$, `shrinkage toward a multifactor model' or $C_{\sf factor}$, `sparse partial correlations' or $C_{\sf sparse}$, and `sparse partial correlations with latent units' or $C_{\sf sparse+latent}$.  First, we demonstrated that, in simulations with known true low-dimensional correlation structures, regularized estimators with matching types of low-dimensional structure generally dominated over the other estimators. We then used cross-validation to determine the most efficient estimator for the population activity of dense groups of neurons in mouse primary visual cortex. We found that $C_{\sf sparse+latent}$ consistently dominated the other estimators.  Estimate produced by $C_{\sf sparse+latent}$ revealed networks of interactions that differed substantially from networks of strongest correlations and depended strongly on the physical distance separating pairs of cells and on the differences in their preferred orientations.

@@

Finally, we consider the fourth estimator $C_{\sf sparse+latent}$, which provides for both common latent factors interacting with all recorded neurons and sparse interactions between the recorded neurons (Fig.~\ref{fig:02}\,D). This estimator has the form \cite{Chandrasekaran:2010,Ma:2013}
\begin{equation}
C_{\sf sparse+latent} = (S - LL^\T)^{-1}
\end{equation}
where, as above, $S$ is a sparse matrix and $L$ is a $d\times p$ matrix of factor loadings. The estimator has two hyperparameters: the number of latent units $d$ and the sparsity of $S$.

@@

In simulation, CV-loss accurately reproduced the differences between the estimators' excess losses, although with greater variability (Figure \ref{fig:02}, Row 5). For each kind of ground truth (Row 2), relative CV-losses were computed with respect to the estimator whose regularization target matched the structure of the respective ground truth: $\ell_{C,C_{\sf diag}}$, $\ell_{C,C_{\sf factor}}$, $\ell_{C,C_{\sf sparse}}$, and $\ell_{C,C_{\sf sparse+latent}}$. Just as with excess loss in Row 4, the means and their standard errors were computed from 30 samples  taken for each ground truth and for each sample size.

@@

In these highly localized populations both direct interactions between cells and common diffuse inputs are likely to contribute to the overall population variability, leading us to hypothesize that regularized estimates capable of accommodating partial correlations between specific pairs of cells (\emph{e.g.}\;$C_{\sf sparse}$ and $C_{\sf sparse+latent}$) would be more efficient than others. At the same time, average correlations were relatively small (Fig.~\ref{fig:03}\,G), to the advantage of estimates biased toward independence (\emph{e.g.}\;$C_{\sf diag}$). Finally, there is the possibility that the correlations structure is best explained by common fluctuations across the entire population, to the advantage of estimates biased toward low-rank correlation structure (\emph{e.g.}\;$C_{\sf factor}$, $C_{\sf factor+lowrank}$).

We found that estimator $C_{\sf sparse+latent}$ was more efficient than the other estimators (Fig.~\ref{fig:04}). The median relative CV-validation loss (Eq.~\ref{eq:rel-cv-loss}) of each of the estimates $C_{\sf 0}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ with respect to $C_{\sf sparse+latent}$ was significantly above zero ($p<0.001$, Wilcoxon signed rank test, for each).

\subsection*{Relationship of correlation structure to circuit architecture}
Having demonstrated that $C_{\sf sparse+latent}$ dominated the other estimators, we examined the elements of the low-dimensional structure revealed by this estimator for individual imaged sites. By its design, the sparse+latent estimator finds the optimal balance between a sparse network of partial correlations and shared common latent units. If this estimate dominates all other estimators evaluated thus far, it seems reasonable to hypothesize that these interaction may reflect underlying physiological interactions more closely than the usual correlations. In particular, the sparse component of the partial correlation matrix suggests direct interactions between pairs of neurons, whereas the low-rank component suggest common fluctuations such as those caused by common inputs or other collective synchronized activity.

We examined the structure of the sparse+latent estimate in individual sites such as the example site first depicted in Fig.\;\ref{fig:03}. In this site, as in others, the regularized estimate of the correlation matrix appeared very similar to the unregularized sample correlation matrix (Fig.~\ref{fig:05}\,A and D). However, the corresponding partial correlation matrices differed dramatically (Fig.~\ref{fig:05}\,B and E). The partial correlation was decomposed into the sparse and low-rank components (Fig.\,\ref{fig:05}\,C). Although correlations were mostly positive, the sparse partial correlations (or `interactions'), although mostly positive, had a much larger proportion of negative values. The sparse component had 82.1\% sparsity (or 17.9\% connectivity), which corresponded to the average node degree (interactions per cell) of 52.5 (Fig.\;\ref{fig:05}\,G). The low-rank component was of rank 17.

In previous studies of the structure of correlations, a sparse network was produced by thresholding the correlations coefficients  above some level deemed significant and examining the network of correlations above that threshold \cite{Golshani:2009,Malmersjo:2013}. There was fairly little overlap between the network of interactions revealed by this thresholding method and those revealed by the sparse+latent estimator (Fig.\;\ref{fig:05}\,F). When thresholded to the same sparsity (82.1\%), only 42\% of cell pairs connected in one network were connected in the other while the average magnitude of such correlations was much lower in the case of the regularized estimator (Fig.\;\ref{fig:05}\,F, H, and I). In particular, many low sample correlations translated into negative interactions in the regularized estimate. Indeed, the absence of a correlation between pairs of cells that both correlate similarly to several of their neighbors should be considered as significant as a high correlation coefficient. Regularized partial correlations reveal such phenomena whereas regular sample correlations cannot.

@@

\begin{figure}[!ht]
    \begin{center}
        %\includegraphics[width=8.3cm]{Figure04.pdf}
    \end{center}
    \caption{{\bf The sparse+latent estimator $C_{\sf sparse+latent}$ outperforms the other estimators on neural data.}
    {\bf A--D.} Histograms of average cross-validation loss differences of the respective estimators $C_{\sf 0}$, $C_{\sf diag}$, $C_{\sf factor}$, and $C_{\sf sparse}$ from $C_{\sf sparse+latent}$. 
    The histograms are based on 31 imaged sites in 24 mice. 
    All medians (red dashed lines) were significantly greater than zero, indicating the dominance of $C_{\sf sparse+latent}$ over the other estimators. 
    The green arrows indicate the results for the site shown in Fig.~\ref{fig:03} and Fig.~\ref{fig:05}
    }
    \label{fig:04}
\end{figure}
